### 2018-03-05 paper/sexSelectionSpacing-ver1.tex
Edit of Descriptive Statistics section

Appendix Table \ref{tab:des_stat1} presents descriptive statistics for
the spells by education level and when the spell began.
There is a substantial number of censored observations.
As an example, for highly educated women who had their first child in the 1995--2006
period, almost half did not have their second child by the time of the survey.
Hence, although about 13,000 women began the second spell 
there are only about 7,000 births to these women.
Censoring becomes even more important for the third and fourth
spells, where around 70 percent of the observations are censored.
Generally, censoring increases with parity and time period.
This reflects a combination of factors: timing of the surveys
relative to the periods of interest, later beginning of childbearing, 
falling fertility, and the longer spells from sex-selective abortions.

The descriptive statistics provide a first indication of how the
sex ratio at birth changes over time and by spell.
For the first spell the sex ratio is very close to the natural
for all education groups and all three time periods.%
\footnote{
There still appears to be some recall error for the group of women without
education for the 1972--1984 period, where 52.3 percent of the children born were boys.
}
For the second spell, all but the highly educated group in the last two
periods have sex ratios in line with the natural sex ratio.
Women with 8 or more years of education have 53.1 and 54.3 percent
boys in the 1985--1994 and 1995--2006 periods, respectively.
This pattern repeats itself for the third spell, except the percentage
boys is higher for the high education group (55.3 and 55.9 for the last
two time periods).
Finally, for the fourth spell the high education group 
had 60 percent boys in the last period, i.e. after the PNDT act was introduced.
Note, however, that for the fourth spell the number of births is substantially
smaller and censoring even more important than for the other spells.

India's population has become progressively more urban.
For the first period, 32 percent of the women entering the first spell lived in urban areas.
This increases to 35 percent for the second period and to 42 percent for the final period.
The population is also substantially better educated.
Women with no education constituted almost 60 percent 
in the first period, but less than thirty percent in the last period.
Correspondingly, in the first period just over twenty percent had 8 or more 
years of education, whereas in the last period it was almost half.
Part of the increase in education is correlated with the increase in urbanization,
but the proportion of better-educated women has increased substantially
in the rural areas as well.
Among the high education group almost 70 percent lived in urban areas
during the first period but this had fallen to less than 60 percent
in the last period.

% The increase in urbanization and education is likely to exert downward pressure
% on fertility and the high censoring rates for the later periods are evidence of this.
% The average number of children born to women by the time they turn 35 illustrate how
% strong the decline in fertility has been.%
% \footnote{
% Figure \ref{fig:fertility} shows the full results.
% }
% Women born in the early 1940s had on average close to 5 children when 
% they reached 35, but women born in the early 1970s had only just over three children.
% The low number of children is especially remarkable because it combines
% all education levels and all areas of residence.
% % Hence, fertility in cities is likely substantially lower.

>>>>

Appendix Table \ref{tab:des_stat1} presents descriptive statistics for
the spells by education level and period they began.
There is a substantial number of censored observations.
As an example, for highly educated women who had their first child in the 2005--2016
period, more than 40\% did not have their second child by the time of the survey.
Censoring becomes even more important for the third and fourth
spells, with more than 70 percent of the observations censored.
The level of censoring also increases with parity and time period,
which reflects a combination of factors: timing of the surveys
relative to the periods of interest, later beginning of childbearing, 
falling fertility, and the increase in spell lengths from 
sex-selective abortions.

The share of urban women in the sample has fallen slightly over the
periods, even though India's population has become progressively more urbanized.
For the first and second periods, over 32\% of the women 
entering the first spell lived in urban areas, falling to 
30\% for the third period and 26\% in the last period.
The most likely explanation is that the age of marriage has increased
faster in urban areas than in rural areas.
Hence, there are relatively more urban women, but fewer show up in the
sample because they are not yet married.

The population has become substantially better educated over time.
Women with no education constituted almost 60 percent 
in the first period, but less than twenty percent in the last period.
Correspondingly, in the first period just over twenty percent had eight or more 
years of education, whereas in the last period more than 60\% did.
These changes are an underestimate of the increase in female
education overall because many of the younger women with more education
have not yet married than therefore are not in the sample.%
\footnote{
Women with eight or more years of education accounted for 65.9\% of
unmarried women in NFHS-3 and 82.8\% of unmarried women in NFHS-4.
See \citet[p 56]{International-Institute-for-Population-Sciences-IIPS2007}
and \citet[p 61]{International-Institute-for-Population-Sciences-IIPS2017}
for more information.
 }

Age at marriage increased over time across all three education groups.
The most substantial increase was for women without any education, where
the average went from below 16 years of age to 18.5.
The smallest change is for the most educated women where the average age
at marriage went up by less than a year---from 19.6 to 20.4---across the 
four periods.
